---
title: "IEA Indicator_MRIP"
author: "Carissa Gervasi"
format: html
editor: visual

---

## Practice IEA GOM indicator - MRIP effort data

### Step 1: pull the data from the online MRIP data query tool.

The MRIP data we pulled here is effort in angler trips for each GOM state by fishing area, private/rental boats only, 1981-2021

```{r}
MRIP = read.csv("../Data/Raw/MRIP query_effort_81_21_GOM by state_private_by area.csv")
summary(MRIP)
```

### Step 2: Summarize the data

Our goal here is to plot overall fishing effort throughout the GOM over time. 

```{r}
#| message: false
library(dplyr)

MRIP.index = as.data.frame(MRIP %>% 
  group_by(Year) %>% 
  summarize(effort = sum(Angler.Trips)))

MRIP.index$effort_mil = MRIP.index$effort/1000000
head(MRIP.index)
str(MRIP.index)

write.csv(MRIP.index, "../Data/Output/MRIP_index.csv", row.names = F)
```

### Step 3: Plot the index

We want to plot effort in angler hours over time. For simplicity, we will plot effort in million angler hours. 

```{r}
#| message: false
library(ggplot2)

ggplot(MRIP.index, aes(x=Year, y=effort_mil)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  labs(x="Year", y="Recreational effort (million angler trips)") +
  scale_x_continuous(breaks=seq(1980,2022,2)) +
  scale_y_continuous(breaks=seq(10,40,1)) +
  theme(axis.title.x = element_text(vjust=-2),
        axis.title.y = element_text(vjust=5)) + 
  theme(axis.text.x = element_text(angle=45, hjust=1)) + 
  theme(plot.margin = unit(c(0.5,0.8,1,1), "cm"))
```

